
---
title: "Mini-Project 2"
author: 
date: "2019-03-24"
output: 
  html_document:
    code_folding: show
---

```{r, message=FALSE, echo=FALSE}
#load packages----
library(tidyverse)
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r, message=FALSE, echo=FALSE}
#select relevant data columns----
candidates_data<-candidates %>%
  select(cand_id, cand_name, cand_election_yr,cand_party_affiliation,cand_office_state) %>%
  filter(cand_election_yr=="2012")
```

```{r, message=FALSE, echo=FALSE}
#select relevant data columns---- 
contributions_data<-contributions %>%
  select(cand_id,name,transaction_amt)
```

```{r, message=FALSE, echo=FALSE}
#join data tables----
candidate_contribution_join <- candidates_data %>%
  full_join(contributions_data, by="cand_id") 

```


```{r, message=FALSE, echo=FALSE}
#filtering, grouping, summarizing data----
money_states<-candidate_contribution_join %>%
  filter(cand_party_affiliation=="DEM"|cand_party_affiliation=="REP") %>%
  group_by(cand_office_state, cand_party_affiliation) %>%
  summarise(total_amt=sum(abs(transaction_amt),na.rm=TRUE)) %>%
  filter(!cand_office_state %in% c("PR","US","VI","AS","GU","MP","DC"))
```

```{r, message=FALSE, echo=FALSE}
#function to create ratio of republican to democrat----
ratio_function<-function(state_arg) {
  money_states %>%
    filter(cand_office_state == state_arg)%>%
  mutate(total = sum(total_amt),
ratio = (total_amt/total)*100)
}

#map function to all states
map_ratio <- map_df(c("AK","AL", "AR","AZ", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "IA", "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME", "MI", "MN", "MO", "MS", "MT", "NC", "ND", "NE", "NH", "NJ", "NM", "NV", "NY", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VA", "VT", "WA", "WI", "WV", "WY"),ratio_function)
```

```{r, message=FALSE, echo=FALSE}
#reordering dataframe to change order of states in plot----
map_ratio_order <- map_ratio %>%
  arrange(-ratio)%>%
  mutate(rep_percent = ifelse(
    cand_party_affiliation == "REP", 1, 0))%>%
  arrange(-rep_percent)
```

```{r,fig.height=12, fig.width=10, message=FALSE, echo=FALSE}
#plotting dataframe----
ratio_plot <-ggplot(map_ratio_order, aes(x = fct_inorder(cand_office_state), y = ratio, 
          fill = cand_party_affiliation))+
  geom_bar(stat="identity", position="fill")+
  scale_fill_manual(values = c("DEM" = "#0C25D4", "REP" = "#DC143C"), 
    labels = c("DEM" = "Democrat", "REP" = "Republican"))+
  theme_minimal()+
  labs(title = "Money Donated To or Against Candidates in 2012 House Election", subtitle = "Organized by State and Candidate's Party", x= "State", y="Percent of Donations by Candidate's Party", fill="Party Affiliation")+
  coord_flip()+
 geom_hline(yintercept = 0.5, color = "black")+
  scale_y_continuous(labels = c("0%", "25%", "50%", "75%", "100%"))

ratio_plot
```

This data graphic demonstrates the monetary influence on Democratic and Republican candidates in all 50 states in the 2012 House election, by showing the ratio of total donations to (or against) Democratic candidates to total donations to (or against) Republican candidates. In this graphic, the decision to sum all donations related to a candidate, whether these donations were towards the candidate or against the candidate, demonstrates which parties were the focus of campaign donations, whether these donations had a negative or positive influence on that candidate. By showing which candidates were the focus of campaign donations, whether those donations were in favor of or against that candidate, this graphic demonstrates which parties were the target of monetary influence. In order to simplify the graphic, donations to or against independent candidates or third-party candidates were removed; the 100 percent total of donations reflects only donations made to or against candidates in the two major parties.

 In some states, like Vermont and Delaware, the candidates that were the focus of donations were almost entirely Democrats, while in states like Alaska and Wyoming, the candidates that were the focus of donations were almost exclusively Republicans. Many states, including Iowa, Colorado, Florida, and Ohio, had an almost-even split the parties of the candidates that were the focus of donations.

Knowing the legislative outcomes of the 2012 Congressional Elections, it is clear that there is a strong correlation between political donations and campaign outcomes. For example, Democratic candidates within Vermont during the 2012 elections were the focus of overwhelmingly more contributions than their Republican counterparts. Similarly in Alaska, no campaign contributions were donated to, nor donated against any Democratic candidates. Our data shows the amount of money affiliated with Democratic and Republican candidates; this money was either donated to or against each candidate representing their parties. Take, for example, Alaska. Because Alaska is historically a Red state, 100 percent of campaign contributions were associated with the Republican candidate. This means that money only went to and against Republican candidates. 

**What we can conclude from these data is that in states with a strong political affiliation, Democratic or Republican, candidates in that political party will be the primary focus of campaign donations, whether these donations are to or against that candidate.** For example, in Vermont, a state that is historically Democratic, nearly all campaign donations focus on Democratic candidates, whether these donations are in support of or against them. This illustrates that in Vermont, there is more monetary focus overall on Democratic candidates; Republicans, rather than donating to support their own candidates, appear to be more likely to donate against Democratic candidates. This idea can be generalized in most other states that are historically Democratic or Republican; donations focus on the party that tends to win, whether these donations are in support of or against that party.

The data graphic also illustrates how swing states can be identified through party contribution. The states where the focus of donations to each party candidate was near the 50% line were seen as swing states in 2012. The Washington Post named nine swing states in 2012. All of the states mentioned fall very close to the 50% line. Iowa, Ohio, Colorado, Florida all relatively evenly split between money to Democrat and Republican candidates. If a race is seen as contentious or people think that there is a chance a seat may “swing” to the other party, the donations focusing on a candidate of each party become close. Although money is connected to election outcomes, in the case of swing states and party donation the outcome cannot be predicted by looking at the amount in donations a candidate is receives. Though outcome cannot be identified through candidate contributions, the status of swing state can be seen through contributions focusing on Republican and Democratic candidates. 

Word count: `r wordcountaddin::word_count()`